Refinements for open automata

### 2022-07-07 2021-Refinement/paper/refinementarticle.tex
meeting notes

\TODO{1 page}
\TODO{move to the end?}

@@

\subsection{Preliminaries and notations}
Families of values, or equivalently maps will be noted \(\mset{i \mapsto x_i}{i \in I}\), or \(\mset{i \gets x_i}{i \in I}\), or \(x_i^{i \in I}\) depending on what is more convenient. % TODO: introduire le fait que \exists c_j^{j \in J} défini J et {j \mapsto c_j}

%For instance \(\mpar{ax}^{x \in \setR}\) represents a scaling function, \(c^{i \in I}\) is a constant function over \(I\).
%They will be used depending on what is more convenient.
%For instance \(\mbrc{\alpha \mapsto 1, \beta \mapsto 2, \gamma \mapsto 3}\) has no simple generating expression and is better represented with the finite version of first notation.
The disjoint union on set is noted \(\uplus\).
%\footnote{Depending on the context, \(\uplus\) notation either supposes that the sets are disjoint or constraints joined sets to be disjoints if several choices are possible}.
Disjoint union is also used on maps.
There are several ways of ensuring a union is disjoint, we will indifferently either suppose sets are disjoint or rename conflicting object (useful for variables).
%The disjoint union of two maps \(\varphi: I \to X\) and \(\psi: J \to Y\) with \(I \cap J = \emptyset\) is noted \(\varphi \uplus \psi\) and has the following signature \(I \uplus J \to X \cup Y\).
In a formula, a quantifier followed by a finite set will be used as a shorthand for the quantification on every variable in the set:
\(\forall \mbrc{a_1, \dots, a_n}, \exists \mbrc{b_1, \dots, b_m}, P\) means \(\forall a_1, \dots, \forall a_n, \exists b_1, \dots, \exists b_m, P\).

@@

\begin{figure}[h]
 \centering
   \includegraphics[width=.5\textwidth]{Figures/counter.pdf}\hfill 
   \includegraphics[width=.3\textwidth]{Figures/compose_count_db.pdf}
   \caption{Open Automaton of \texttt{counter} process (left).  Its role is to get a target, then count ticks until that target is reached, then emits an action with the elapsed time and restart. And Open Automation of  the composition \texttt{database}[\texttt{counter}] (right).\label{Fig:counter}}
\end{figure}
\paragraph{Relations between open automata}
A relation (bisimulation, refinement, etc.) between OA requires to compare their states. To do so we will suppose that the variables of the two OA are disjoint (a renaming of variables may have to be applied before comparing OA states).
\begin{definition}[Relation on OA states] Suppose $V_1$ and $V_2$ are disjoint.
A relation on states of \(\OAg[1]\) and \(\OAg[2]\) is a function \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\).
\end{definition}
The idea is that two states are related if the variables they refer to verify a certain formula. Additionally, we may check that initial states of the automata are related by checking that: \(\sigma_{01} \uplus \sigma_{022} \vdash R\mpar{s_{01}, s_{02}}\).

@@

TODO OR NOT?

@@

>>>>

@@

\subsection{Preliminaries and notations}
Families of values, or equivalently maps will be noted \(x_i^{i \in I}\), \(\mset{i \mapsto x_i}{i \in I}\), or \(\mset{i \gets x_i}{i \in I}\), depending on what is more convenient. % TODO: introduire le fait que \exists c_j^{j \in J} défini J et {j \mapsto c_j}
%For instance \(\mpar{ax}^{x \in \setR}\) represents a scaling function, \(c^{i \in I}\) is a constant function over \(I\).
%They will be used depending on what is more convenient.
%For instance \(\mbrc{\alpha \mapsto 1, \beta \mapsto 2, \gamma \mapsto 3}\) has no simple generating expression and is better represented with the finite version of first notation.
The disjoint union on set is noted \(\uplus\).
%\footnote{Depending on the context, \(\uplus\) notation either supposes that the sets are disjoint or constraints joined sets to be disjoints if several choices are possible}.
Disjoint union is also used on maps.
There are several ways of ensuring a union is disjoint, we will indifferently either suppose sets are disjoint or rename conflicting object (useful for variables).
%The disjoint union of two maps \(\varphi: I \to X\) and \(\psi: J \to Y\) with \(I \cap J = \emptyset\) is noted \(\varphi \uplus \psi\) and has the following signature \(I \uplus J \to X \cup Y\).
In a formula, a quantifier followed by a finite set will be used as a shorthand for the quantification on every variable in the set:
\(\forall \mbrc{a_1, \dots, a_n}, \exists \mbrc{b_1, \dots, b_m}, P\) means \(\forall a_1, \dots, \forall a_n, \exists b_1, \dots, \exists b_m, P\).

@@

\begin{figure}[h]
 \centering
   \includegraphics[width=.5\textwidth]{Figures/counter.pdf}\hfill 
   \includegraphics[width=.3\textwidth]{Figures/compose_count_db.pdf}
   \caption{Open Automaton of \texttt{counter} process (left).  Its role is to get a target, then count ticks until that target is reached, then emits an action with the elapsed time and restart. And Open Automation of  the composition \texttt{database}[\texttt{counter/cnt}] (right).\label{Fig:counter}}
\end{figure}
\paragraph{Relations between open automata}
A relation (bisimulation, refinement, etc.) between OA requires to compare their states. To do so we will suppose that the variables of the two OA are disjoint (a renaming of variables may have to be applied before comparing OA states).
\begin{definition}[Relation on OA states] Suppose $V_1$ and $V_2$ are disjoint.
A relation on states of \(\OAg[1]\) and \(\OAg[2]\) is a function \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\).
\end{definition}
The idea is that two states are related if the variables they refer to verify a certain formula. Additionally, we may check that initial states of the automata are related by checking that: \(\sigma_{01} \uplus \sigma_{022} \vdash R\mpar{s_{01}, s_{02}}\).

@@

Bisimulation between OA was defined in~\cite{fh-bisim}. Instead of recalling its definition we show below the principles of this bisimulation.

\begin{tikzpicture}
  \node[state,inner sep=5pt,minimum size=10pt] (1) {$s$};
  \node[state,inner sep=5pt,minimum size=10pt, below right of=1,right=80pt] (2) {$t$};
  \node[state,inner sep=5pt,minimum size=10pt, above right of=1,right=80pt] (3) {$t'$};

  \draw (1) edge[align=center] node[below]{\OT{}{}{\alpha(x)}{h\mapsto\beta(x)}{x<0}{y\leftarrow -x}}  (2);
  \draw (1) edge[align=center] node[above]{\OT{}{}{\alpha(x)}{h\mapsto\beta(x)}{x\geq 0}{y\leftarrow x}} (3);
\end{tikzpicture}
\begin{tikzpicture}
  \node[state,inner sep=5pt,minimum size=10pt] (1) {$s_2$};
  \node[state,inner sep=5pt,minimum size=10pt,  right of=1,right=90pt] (2) {$t_2$};

  \draw (1) edge[align=center] node[below]{\OT{}{}{\alpha(x)}{h\mapsto\beta(x)}{x<0}{y\leftarrow -x}}  (2);
  \draw (1) edge[align=center] node[above]{\OT{}{}{\alpha(x)}{h\mapsto\beta(x)}{x\geq 0}{y\leftarrow x}} (3);
\end{tikzpicture}

@@

\TODO{1 page}
